### 2017-04-13 paper/oopsla17-submission/intro2.tex
pass over intro2

Our \emph{first contribution} (\autoref{sec:overview})
is a set of more than 4,500 labeled programs,
gives us an accurate ground truth of
the kinds of errors and the (locations
of the) fixes that novices make in
practice.
%
We obtain this set by observing that
software development is an iterative
process; programmers eventually
fix their own ill-typed programs,
perhaps after multiple incorrect
exploratory attempts.
%
To exploit this observation we instrumented
the \ocaml compiler to collect fine-grained
traces of student interactions over two instances
of an undergraduate Programming Languages course
%
\footnote{
% \begin{anonsuppress}
% UC San Diego (IRB \#140608).
% \end{anonsuppress}
% \begin{noanonsuppress}
at AUTHOR's INSTITUTION (IRB HIDDEN).
% \end{noanonsuppress}
}
%
We then post-process the resulting time-series
of programs submitted to the \ocaml compiler into
a set of pairs of ill-typed programs and their
subsequent \emph{fixes}, the first (type-)correct
program in the trace suffix.
%
Finally, we label the changed locations by computing a
\emph{tree-diff} between the two terms to find
the exact sub-terms that were changed in the fix.

\mypara{Challenge 2: Modeling Programs as Vectors}
%
Modern supervised learning algorithms work on sets of
\emph{feature vectors}: real-valued points in an
$n$-dimensional space. While there are standard
techniques for computing such
vectors for documents, images, and sound (respectively
word-counts, pixel-values, and frequencies), it has
hitherto been unclear how to precisely model program
structure as numeric data.

Our \emph{second contribution} (\autoref{sec:learning})
solves this problem with a new representation called
a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
%
(1)~each program is represented by the \emph{bag}
    or multiset of (sub-)terms that appears inside
    it, and
%
(2)~each (sub-)term $t$ is \emph{abstracted} as
    a vector $[f_1(t), \ldots, f_n(t)]$ where
    each $f_i$ is a \emph{feature abstraction}
    that maps a term to a numeric value.
%
We can model the error labels with a simple feature
abstraction (\eg is-a-diff-subterm).
%
We can even account for \emph{contextual} information like
the features of the parent- and children- terms by
concatenating the feature vectors of each term with those
of its parent and children.

\mypara{Challenge 3: Training Precise Classifiers}
%
Finally, the last and most important challenge is to
use the labeled BOAT data to train classifiers that
are capable of \emph{precisely} pinpointing the error
locations in real programs.
%
The key here is to find the right set of feature
abstractions and classification algorithms, that
lead to precise predictions.
%
Fortunately, our BOAT model allows us a great deal of
latitude in our choice of features.
%
We can use abstraction functions to capture different
aspects of a term ranging from
%
syntactic features (\eg is-a-data constructor, is-a-literal,
is-an-arithmetic-operation, is-a-function-application \etc),
%
to semantic features captured by the type system (\eg is-a-list,
is-an-integer, is-a-function \etc).

Our \emph{third contribution} (\autoref{sec:evaluation})
is a systematic evaluation of our data-driven approach
using different classes of features like the above, and
with three different classification algorithms: linear
classifiers, decision forests and deep neural networks.
%
Our evaluation lets us identify the best features and
classifier and empirically characterize the importance
of different features for error localization.
%
In particular, we find that while machine learning
on its own performs slightly worse than existing
purely constraint based approaches (\eg \ocaml, \sherrloc),
once we augment the data with feature abstraction
corresponding to the \emph{type error slice} \cite{Tip2001-qp},
the resulting classifier significantly outperforms
state-of-the-art by 18 percentage points.

>>>>

Our \emph{first contribution} (\autoref{sec:overview})
is a set of more than 4,500 labeled programs,
gives us an accurate ground truth of
the kinds of errors and the (locations
of the) fixes that novices make in
practice.
%
We obtain this set by observing that
software development is an iterative
process; programmers eventually
fix their own ill-typed programs,
perhaps after multiple incorrect
exploratory attempts.
%
To exploit this observation we instrumented
the \ocaml compiler to collect fine-grained
traces of student interactions over two instances
of an undergraduate Programming Languages
course.\footnote{
% \begin{anonsuppress}
% UC San Diego (IRB \#140608).
% \end{anonsuppress}
% \begin{noanonsuppress}
at AUTHOR's INSTITUTION (IRB HIDDEN).
% \end{noanonsuppress}
}
%
We then post-process the resulting time-series
of programs submitted to the \ocaml compiler into
a set of pairs of ill-typed programs and their
subsequent \emph{fixes}, the first (type-)~correct
program in the trace suffix.
%
Finally, we label the changed locations by computing a
\emph{tree-diff} between the two terms to find
the exact sub-terms that were changed in the fix.

\mypara{Challenge 2: Modeling Programs as Vectors}
%
Modern supervised learning algorithms work on %sets of
\emph{feature vectors}: real-valued points in an
$n$-dimensional space. While there are standard
techniques for computing such
vectors for documents, images, and sound (respectively
word-counts, pixel-values, and frequencies),
% it has
% hitherto been unclear how to precisely model program
% structure as numeric data.
there are no similarly standard representations for
programs.

Our \emph{second contribution} (\autoref{sec:learning})
solves this problem with a simple, yet expressive, representation called
a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
%
(1)~each program is represented by the \emph{bag}
    or multiset of (sub-)~terms that appears inside
    it; and
%
(2)~each (sub-)~term $t$ is \emph{abstracted} as
    a vector $[f_1(t), \ldots, f_n(t)]$, where
    each $f_i$ is a \emph{feature abstraction}
    that maps a term to a numeric value.
%
We can model the error labels with a simple feature
abstraction (\eg does-change-in-fix).
%
We can even account for \emph{contextual} information like
the features of the parent- and child- terms by
concatenating the feature vectors of each term with those
of its parent and children.
%
We have found this representation to be particularly
convenient as it gives us flexibility
in modeling the syntactic and semantic structure of programs
while retaining compatibility with off-the-shelf classifiers,
in contrast to, \eg, \citet{Raychev2015-jg}, who had to develop
their own variants of classifiers to obtain their results.

\mypara{Challenge 3: Training Precise Classifiers}
%
Finally, the last and most important challenge is to
use the labeled BOAT data to train classifiers that
are capable of \emph{precisely} pinpointing the error
locations in real programs.
%
The key here is to find the right set of feature
abstractions and classification algorithms, that
lead to precise predictions.
%
Fortunately, our BOAT model allows us a great deal of
latitude in our choice of features.
%
We can use abstraction functions to capture different
aspects of a term ranging from
%
syntactic features (\eg is-a-data constructor, is-a-literal,
is-an-arithmetic-operation, is-a-function-application, \etc),
%
to semantic features captured by the type system (\eg is-a-list,
is-an-integer, is-a-function, \etc).

Our \emph{third contribution} (\autoref{sec:evaluation})
is a systematic evaluation of our data-driven approach
using different classes of features like the above, and
with three different classification algorithms: logistic
regression, decision trees, and neural networks.
%
Our evaluation lets us identify the best features and
classifier and empirically characterize the importance
of different features for error localization.
%
In particular, we find that while machine learning
over syntactic features of each term in isolation
performs worse than existing
purely constraint based approaches (\eg \ocaml, \sherrloc),
augmenting the data with a single feature corresponding to
the \emph{type error slice} \citep{Tip2001-qp} brings our
classifiers up to par with the state of the art,
and further augmenting the data with \emph{contextual}
features allows our classifiers to outperform
the state of the art by 18 percentage points.
% In particular, we find that while machine learning
% on its own performs slightly worse than existing
% purely constraint based approaches (\eg \ocaml, \sherrloc),
% once we augment the data with feature abstraction
% corresponding to the \emph{type error slice} \citep{Tip2001-qp},
% the resulting classifier significantly outperforms the
% state of the art by 18 percentage points.
